Prediction of Weight Lifting Style using Accelerometer Data
Introduction



========================================================

With the availability of low cost accelerometers, there are many opportunities to measure human activities. One application of this is [measuring the proper form of weight lifting][wle]. In this paper we examine whether we can determine the weight lifting form using the accelerometer data collected.

```{r}
raw_training <- read.csv('C:/dung/dataScience/machineLearning/pml-training.csv')
raw_testing <- read.csv('C:/dung/dataScience/machineLearning/pml-testing.csv')
library(caret)
set.seed(1234)
trainingIndex <- createDataPartition(raw_training$classe, list=FALSE, p=.9)
training = raw_training[trainingIndex,]
testing = raw_training[-trainingIndex,]
```
Remove indicator with Zero Variance

```{r}
library(caret)
nzv <- nearZeroVar(training)

training <- training[-nzv]
testing <- testing[-nzv]
raw_testing <- raw_testing[-nzv]
```
Filter columns to only include numeric features and outcome. Integer and other non-numeric features can be trained to reliably predict values in the training file provided, but when used to predict values in the testing set provided, they lead to misclassifications.

```{r}
library (RANN)
num_features_idx = which(lapply(training,class) %in% c('numeric')  )
preModel <- preProcess(training[,num_features_idx], method=c('knnImpute'))

ptraining <- cbind(training$classe, predict(preModel, training[,num_features_idx]))
ptesting <- cbind(testing$classe, predict(preModel, testing[,num_features_idx]))
prtesting <- predict(preModel, raw_testing[,num_features_idx])

#Fix Label on classe
names(ptraining)[1] <- 'classe'
names(ptesting)[1] <- 'classe'
```

Model

We can build a random forest model using the numerical variables provided. 

```{r}
library(randomForest)
rf_model  <- randomForest(classe ~ ., ptraining, ntree=500, mtry=32)

```
Cross validation
```{r}
training_pred <- predict(rf_model, ptraining) 
print(confusionMatrix(training_pred, ptraining$classe))
```

OUt of sample accuracy

```{r}
testing_pred <- predict(rf_model, ptesting) 
```

Confusion matrix

```{r}
print(confusionMatrix(testing_pred, ptesting$classe))

```
The cross validation accuracy is greater than 99%, which should be sufficient for predicting the twenty test observations. Based on the lower bound of the confidence interval we would expect to achieve a 98.7% classification accuracy on new data provided.

Test Set Prediction Results

Applying this model to the test data provided yields 100% classification accuracy on the twenty test observations.

```{r}
answers <- predict(rf_model, prtesting) 
answers
```


Conclusion

We are able to provide very good prediction of weight lifting style as measured with accelerometers.

